\subsection*{Proposal Summary:}
% Proposal summary (identical to the abstract from the online proposal submission forms, section 1). 

% The abstract (summary) should, at a glance, provide the reader with a clear understanding of the objectives of the research proposal and how they will be achieved. The abstract will be used as the short description of your research proposal in the evaluation process and in communications to contact in particular potential  independent external experts and/or to inform the Commission and/or the programme management committees and/or relevant national funding agencies. It must therefore be short and precise and should not contain confidential information. 

% Please use plain typed text, avoiding formulae and other special characters. The abstract must be written in English. There is a limit of 2000 characters (spaces and line breaks included).

% 

There are fundamental differences between the standard scientific method and the method of mathematical research. Broadly speaking, the standard scientific method is founded upon observations; these observations  are eventually used to formulate hypotheses, and those are then confirmed or rejected based on carefully designed experiments. 

In contrast to the natural sciences, 
there is frequently a paucity of initial data upon which hypotheses can be formulated in many mathematical disciplines. 
Quite often experimental methodology is absent, making it challenging to validate conjectures through empirical evidence. Mathematicians are so used to this way of conducting research that the possibility of using experimental data is often overlooked even when available or not too difficult to obtain. This project aims to change this reality in the realm of abstract polytopes.

Abstract polytopes (APs) are combinatorial objects that generalise polygons, polyhedra and other geometrical objects. The study of APs comes hand in hand with the study of their symmetries. When working with symmetric objects, a natural question is: how symmetrical can an object be? More precisely, given certain symmetry conditions, is there such an object that satisfies these conditions? What structural conditions are imposed on such an object by the symmetry constraints? In this proposal, we address these questions for APs and related objects.  

We propose a data-driven approach with a two-way flow of knowledge: with the aim of generating new datasets of highly symmetric APs and related objects, we will develop new theoretical results and methods, enabling us to devise new algorithms and combinatorial representations of APs. The algorithms will then be carefully implemented and executed.
The obtained datasets will then be analysed to find interesting patterns, suggesting new conjectures and proposing new directions for further research.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Explain and justify the cross-panel or cross domain nature of your proposal, if a secondary panel is indicated in the online proposal submission forms. There is a limit of 1000 characters (spaces and line breaks included).